Cannibalism among the supermassive black holes

### 2018-12-11 paper.tex
Corrected Lx equations Black Hole Luminosity section
Minor changes; added t_fric equation

\subsection{Merger tree Data}
The merger tree data that was used as the input to our simulations is the result of previous work by \citet{2011ApJ...741...99C, 2011ApJ...742L..33C, 2012ApJ...753...17C, 2012ApJ...748..121C, 2013ApJ...770..139C} (hereafter Cen11), \citet{2012MNRAS.425..641L} (hereafter Lackner12), and \citet{2015ApJ...799..178K} (hereafter Kulier15).  A brief synopsis of the work that led to our input data is described below.

Cen11 studied the "cosmic downsizing" effect (e.g., \citet{1996AJ....112..839C}) using high-resolution large-scale hydrodynamic galaxy formation simulations.  His work produced physical parameters for galaxies such as position, velocity, total mass, stellar mass, gas mass, mean formation time, mean stellar metallicity, mean gas metallicity, SFR, luminosities, etc.

Lackner12, using Cen11's work as a basis, created galactic merger trees that compared the properties of in situ and accreted stellar mass of galaxies at redshift snapshots between $4 > z > 0$.

Kulier15 furthered the merger tree data produced by Lackner12 by estimating the evolution of the SMBH population, based on assumptions about the relationship between the properties of SMBHs and those of their host galaxies.

@@

The merger tree data from Lackner12 and Kulier15 provided 1,830 galaxies in total.  However, not all the galaxies were suitable for our simulations.  We placed further requirements as follows:
\begin{itemize}
\item The galaxies had to exist through the entire simulation ($4 > z > 0$).  If they merged with other galaxies, they had to have been the "surviving" galaxy at each merger.
\item They had to have accumulated orbiting black holes by $z = 0$.
\end{itemize}

@@

The following cosmological parameters were used in our simulations, consistent with both Lackner12 and Kulier15:   $\Omega_M = 0.28$, $\Omega_b = 0.046$, $\Omega_\Lambda = 0.72$, $\sigma_8 = 0.82$, $H_0 = 100h^{-1}Mpc^{-1} = 70 km s^{-1} Mpc^{-1}$, and $n = 0.96$.

@@

Lackner12, which is the source of our stellar masses, noted that the efficiency of star formation in their simulations, defined as $f_*=M_*/M_{DM}(\Omega_{DM}/\Omega_b)$, was approximately 0.6.  Compared with the expected range of $0.10 \lessapprox f_* \lessapprox 0.15$ that they referenced from \citet{2012ApJ...746...95L}, their stellar masses were a factor of roughly 4 times greater.

We also compared the stellar masses from Lackner12 to  observational and abundance matching data from \citet{2018AstL...44....8K}. Figure \ref{fig:stellar1} is a partial, approximate, replication of their Figure 11 (abundance matching line).  We rescaled our stellar masses at $z=0$ in a lognormal manner about the line in order to match the observed data.  Table \ref{table:gal_char} shows the original and rescaled stellar masses.

@@

\subsubsection{Galaxy background potential} \label{Galaxy background potential}
For our simulations, the Stone-Ostriker profile (\citet{2015ApJ...806L..28S}), which is a three-parameter potential-density pair, whose quantities such as density, potential, and binding energy can be written in closed form.  It is essentially an analytic form of a finite, cored isothermal mass distribution:
\begin{equation} \label{jerry}
\rho(r) = \frac{\rho_c}{(1+r^2/r_{c}^2)(1+r^2/r_{h}^2)}
\end{equation}
Here $\rho_c$ is the central density, $r_c$ is the core radius, and $r_h$ is the outer halo radius.

In order to parametrize this profile, $r_c$ was given an initial value of 100 pc, which is reasonable for a cored massive system.  To calculate $r_h$, we began by using the stellar mass to calculate the velocity dispersion at the galaxy's effective radius.  In Kulier12, the galaxy's effective radius $R_{e}$ and velocity dispersion $\sigma(R_e)$ at the effective radius are:
\begin{equation} \label{re}
R_{e} = 2.5 kpc\left(\frac{M_*}{10^{11}M_{\odot}}\right)^{0.73}(1+z)^{-0.98}
\end{equation}
\begin{equation} \label{sig}
\sigma(R_{e}) = 190km/s\left(\frac{M_{*}}{10^{11}M_{\odot}}\right)^{0.2}(1+z)^{0.47}
\end{equation}

@@

At each iteration in the code, $r_h$ is updated according to the stellar mass and Equations \ref{re} and \ref{sig}.  The core radius, $r_c$, is recalculated only if an orbiting black hole gets within $r_c$.  The work done by diffusion, as described in Section \ref{psd} is calculated, the total potential energy and dynamical friction are updated, and $r_c$ is solved for from Equation 8 in Stone-Ostriker.

@@

\subsubsection{Gravitational wave recoils}
The code \textsc{AR-Chain} includes post-Newtonian terms up to order 2.5. The SMBHs can therefore merge via gravitational wave emission. We include gravitational wave recoils following the prescription outlined in Kulier15, which is based on the fitting formula by \citet{2012PhRvD..85h4015L}.  We assume that a merger will be inevitable when the separation between two SBHs becomes smaller than 1.0 Schwarzschild radius. At the moment of the merger, we assume that the spin vectors of the two SBHs are randomly aligned.

@@

\subsection{Simulation setup}
Using the merger tree data, orbiting black holes were injected into its host galaxy at redshift z, at a distance from the galactic center of $R_{e}$ (Eqn. \ref{re}) following galaxy mergers.  Their initial velocity was arbitrarily chosen to be circular, $v_c = \sqrt{M(R_e)/R_e}$, with $v_x$, $v_y$, and $v_z$ randomly chosen.  Note: we only inject the BHs into the simulation that have a $t_{fric}$ smaller than 100 times the Hubble time.

\subsection{Black Hole Luminosity}
The x-ray temperature versus sigma relation from Table 2 of  \citet{2018ApJ...857...32B} was fitted according to:
\begin{equation}
    T_x = 10^{(2.50*log \sigma - 6.06)}
\end{equation}
Central gas density, ${\rho}_0$ was also fitted from Table 3 of \citet{2018ApJ...857...32B} according to:
\begin{equation}
    \rho_0 = 10^{(0.6log(\sigma) - 25)}
\end{equation}
The core radius of the gas model, ${r_c}$ is:
\begin{equation}
    r_c = 10^{(15.09log(\sigma) - 35.77)}
\end{equation}
Local gas density ${\rho}$ is obtained from the isothermal beta model:
\begin{equation}
    \rho = \frac{\rho_0}{(1+r^2/r_c^2)^{1.5}}
\end{equation}
Gas sound speed, ${ss}_g$, was obtained from:
\begin{equation}
    {ss}_g = \sqrt{1.67T_x/m_p},
\end{equation}
where $m_p$ is the proton mass.

The black hole mass accretion rate is approximated by:
\begin{equation}
    \dot{M}_{acc} = \frac{4{\pi}G^2M^2\rho}{({ss_g}^2+v^2)^{3/2}},
\end{equation}
where M is black hole mass, G is the gravitational constant, and v is the black hole velocity.

Finally, the black hole luminosity is calculated from:
\begin{equation}
    L = 0.1\dot{M}_{acc}{ss_g}^2
\end{equation}

>>>>

\subsection{Merger tree Data}
The merger tree data that was used as the input to our simulations is the result of previous work by \citet{2011ApJ...741...99C, 2011ApJ...742L..33C, 2012ApJ...753...17C, 2012ApJ...748..121C, 2013ApJ...770..139C} (hereafter Cen(+11)), \citet{2012MNRAS.425..641L} (hereafter Lackner(+12)), and \citet{2015ApJ...799..178K} (hereafter Kulier(+15)).  A brief synopsis of the work that led to our input data is described below.

Cen(+11) studied the "cosmic downsizing" effect (e.g., \citet{1996AJ....112..839C}) using high-resolution large-scale hydrodynamic galaxy formation simulations.  His work produced physical parameters for galaxies such as position, velocity, total mass, stellar mass, gas mass, mean formation time, mean stellar metallicity, mean gas metallicity, SFR, luminosities, etc.

Lackner(+12), using Cen(+11)'s work as a basis, created galactic merger trees that compared the properties of in situ and accreted stellar mass of galaxies at redshift snapshots between $4 > z > 0$.

Kulier(+15) furthered the merger tree data produced by Lackner(+12) by estimating the evolution of the SMBH population, based on assumptions about the relationship between the properties of SMBHs and those of their host galaxies.

@@

The merger tree data from Lackner(+12) and Kulier(+15) provided 1,830 galaxies in total.  However, not all the galaxies were suitable for our simulations.  We placed further requirements as follows:
\begin{itemize}
\item The galaxies had to exist through the entire simulation ($4 > z > 0$).  If they merged with other galaxies, they had to have been the "surviving" galaxy at each merger.
\item They had to have accumulated orbiting black holes by $z = 0$.
\end{itemize}

@@

The following cosmological parameters were used in our simulations, consistent with both Lackner(+12) and Kulier(+15):   $\Omega_M = 0.28$, $\Omega_b = 0.046$, $\Omega_\Lambda = 0.72$, $\sigma_8 = 0.82$, $H_0 = 100h^{-1}Mpc^{-1} = 70 km s^{-1} Mpc^{-1}$, and $n = 0.96$.

@@

Lackner(+12), which is the source of our stellar masses, noted that the efficiency of star formation in their simulations, defined as $f_*=M_*/M_{DM}(\Omega_{DM}/\Omega_b)$, was approximately 0.6.  Compared with the expected range of $0.10 \lessapprox f_* \lessapprox 0.15$ that they referenced from \citet{2012ApJ...746...95L}, their stellar masses were a factor of roughly 4 times greater.

We also compared the stellar masses from Lackner(+12) to  observational and abundance matching data from \citet{2018AstL...44....8K}. Figure \ref{fig:stellar1} is a partial, approximate, replication of their Figure 11 (abundance matching line).  We rescaled our stellar masses at $z=0$ in a lognormal manner about the line in order to match the observed data.  Table \ref{table:gal_char} shows the original and rescaled stellar masses.

@@

\subsubsection{Galaxy background potential} \label{Galaxy background potential}
In order to model the dark matter halo distribution in our simulations, we used the Stone-Ostriker profile (\citet{2015ApJ...806L..28S}), which is a three-parameter potential-density pair.  Quantities such as density, potential, and binding energy can be written in closed form.  It is essentially an analytic form of a finite, cored isothermal mass distribution:
\begin{equation} \label{jerry}
\rho(r) = \frac{\rho_c}{(1+r^2/r_{c}^2)(1+r^2/r_{h}^2)}
\end{equation}
Here $\rho_c$ is the central density, $r_c$ is the core radius, and $r_h$ is the outer halo radius.

In order to parametrize this profile, $r_c$ was given an initial value of 100 pc, which is reasonable for a cored massive system.  To calculate $r_h$, we began by using the stellar mass to calculate the velocity dispersion at the galaxy's effective radius.  In Kulier(+15), the galaxy's effective radius $R_{e}$ and velocity dispersion $\sigma(R_e)$ at the effective radius are:
\begin{equation} \label{re}
R_{e} = 2.5 kpc\left(\frac{M_*}{10^{11}M_{\odot}}\right)^{0.73}(1+z)^{-0.98}
\end{equation}
\begin{equation} \label{sig}
\sigma(R_{e}) = 190km/s\left(\frac{M_{*}}{10^{11}M_{\odot}}\right)^{0.2}(1+z)^{0.47}
\end{equation}

@@

At each iteration in the code, $r_h$ is updated according to the stellar mass and Equations \ref{re} and \ref{sig}.  The core radius, $r_c$, is recalculated only if an orbiting black hole enters within $r_c$.  The work done by diffusion, as described in Section \ref{psd} is calculated, the total potential energy and dynamical friction are updated, and $r_c$ is solved for from Equation 8 in Stone-Ostriker.

@@

\subsubsection{Gravitational wave recoils}
The code \textsc{AR-Chain} includes post-Newtonian terms up to order 2.5. The SMBHs can therefore merge via gravitational wave emission. We include gravitational wave recoils following the prescription outlined in Kulier(+15), which is based on the fitting formula by \citet{2012PhRvD..85h4015L}.  We assume that a merger will be inevitable when the separation between two SBHs becomes smaller than 1.0 Schwarzschild radius. At the moment of the merger, we assume that the spin vectors of the two SBHs are randomly aligned.

@@

\subsection{Simulation setup}
For each galaxy in the simulations, we only used black holes that would theoretically have merged with the center of the host galaxy within a dynamical friction time, $t_{fric}$, of less than 100 times a Hubble time, with $t_{fric}$ defined from \citet{2008gady.book.....B} as:
\begin{equation}\label{tfric}
    t_{fric} = \frac{19}{6}Gyr\left(\frac{R_e}{5kpc}\right)^2\frac{\sigma_{eff}}{200km\,s^{-1}}\frac{10^8\,M_{\odot}}{M_{bh}}
\end{equation}
The orbiting black holes were injected into their host galaxy at redshift z, at a distance from the galactic center of $R_{e}$ (Eqn. \ref{re}) following galaxy mergers.  Their initial velocity was arbitrarily chosen to be circular, $v_c = \sqrt{M(R_e)/R_e}$, with $v_x$, $v_y$, and $v_z$ randomly chosen.

\subsection{Black Hole Luminosity}
The x-ray temperature versus sigma relation from Table 2 of  \citet{2018ApJ...857...32B} was fitted according to:
\begin{equation}
    T_x = 10^{(2.50*log \sigma - 6.06)}
\end{equation}
Central gas density, ${\rho}_0$ was also fitted from Table 3 of \citet{2018ApJ...857...32B} according to:
\begin{equation}
    \rho_0 = 10^{(0.6log(\sigma) - 25)}
\end{equation}
The core radius of the gas model, ${r_c}$ is:
\begin{equation}
    r_c = 10^{(15.09log(\sigma) - 35.77)}
\end{equation}
Local gas density ${\rho}$ is obtained from the isothermal beta model:
\begin{equation}
    \rho = \frac{\rho_0}{(1+r^2/r_c^2)^{1.5}}
\end{equation}
Gas sound speed, $c_s$, was obtained from:
\begin{equation}
    c_s = \sqrt{1.67T_x/m_p},
\end{equation}
where $m_p$ is the proton mass.

The black hole mass accretion rate is approximated by:
\begin{equation}
    \dot{M}_{acc} = \frac{4{\pi}G^2M^2\rho}{(c_s^2+v^2)^{3/2}},
\end{equation}
where M is black hole mass, G is the gravitational constant, and v is the black hole velocity.

Finally, the black hole xray luminosity is calculated from:
\begin{equation}
    L_x = 0.01\dot{M}_{acc}c^2
\end{equation}
